\section{Bold, italics and underlining}

Next, we will now look at some text formatting commands:

\begin{itemize}
    \item \textbf{Bold}: bold text in LaTeX is typeset using the \verb|\textbf{...}| command.
    \item \textit{Italics}: italicised text is produced using the \verb|\textit{...}| command.
    \item \underline{Underline}: to underline text use the \verb|\underline{...}| command.
\end{itemize}

The next example demonstrates these commands:

\begin{tcolorbox}
\begin{verbatim}
    Some of the \textbf{greatest}
    discoveries in \underline{science} 
    were made by \textbf{\textit{accident}}.
\end{verbatim}
\end{tcolorbox}

This example produces the following output:

\begin{mdframed}
\-\hspace{20pt}Some of the \textbf{greatest} discoveries in \underline{science} were made by \textbf{\textit{accident}}.
\end{mdframed}

Another very useful command is \verb|\emph{|\emph{argument}\verb|}|, whose effect on its \emph{argument} depends on the context. Inside normal text, the emphasized text is italicized, but this behaviour is reversed if used inside an italicized text—see the next example:

\begin{tcolorbox}
\begin{verbatim}
    Some of the greatest \emph{discoveries} in science were made 
    by accident.

    \textit{Some of the greatest \emph{discoveries} in science were 
    made by accident.}

    \textbf{Some of the greatest \emph{discoveries} in science were 
    made by accident.}
\end{verbatim}
\end{tcolorbox}

This example produces the following output:

\begin{mdframed}
    Some of the greatest \emph{discoveries} in science 
    were made by accident.

    \textit{Some of the greatest \emph{discoveries} 
    in science were made by accident.}

    \textbf{Some of the greatest \emph{discoveries} 
    in science were made by accident.}
\end{mdframed}

\begin{itemize}
    \item \textbf{Note:} some packages, such as Beamer, change the behaviour of the \verb|\emph| command.
\end{itemize}